---
title: 'Data 607: Week 9 TidyVerse/GitHub'
author: Nakesha Fray
output: html_document
date: "2024-11-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Introdution

For this week's assignment, I selected a dataset from fivethirtyeight to create a programming sample “vignette" which will demonstrates how to use one or more of the TidyVerse packages.

The dataset I chose was about weather checking from the article: Where People Go To Check The Weather by Walt Hickey: https://fivethirtyeight.com/features/weather-forecast-news-app-habits/

## Package: readr
This is a tidyverse package that reads in csv. files into R. In this step, I am using readr to pull in the csv. raw file from GitHub.

```{r readr}
weather <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/refs/heads/master/weather-check/weather-check.csv")

glimpse(weather)
```

This dataset has 928 observations and 9 variables. Each observation is a respondent. 

## Package: dplyr

Dplyr has many functions, but I will be using it to rename the columns to something shorter and easier to refer to later on and then select to create a ggplot.

```{r rename}

weather2 <- rename(weather, respondent_id = RespondentID,
                check_weather = Do.you.typically.check.a.daily.weather.report.,
                how_check_weather = How.do.you.typically.check.the.weather.,
                weather_app_web = A.specific.website.or.app..please.provide.the.answer.,
                likert_check_weather = If.you.had.a.smartwatch..like.the.soon.to.be.released.Apple.Watch...how.likely.or.unlikely.would.you.be.to.check.the.weather.on.that.device.,
       gender = What.is.your.gender.,
       household_income = How.much.total.combined.money.did.all.members.of.your.HOUSEHOLD.earn.last.year.,
       us_region = US.Region                   
       )

weather2[weather2 == "-"] <- NA
```

```{r select}
weather3 <- weather2 %>%
  select(respondent_id, how_check_weather)
```

## Package: tidy

Tidy also has many functions but I will be using it to drop NAs.

```{r tidy}
weather4 <- weather3 %>% drop_na()
```

## Package: dplyr

Here we will go back to dplyr to use group_by and summarise to create a count of the variable: how people check the weather.

```{r groupby_and_summarise}

weather4$how_check_weather[weather4$how_check_weather == "A specific website or app (please provide the answer)"] <- "Other"

weather5 <- weather4 %>%
   group_by(how_check_weather) %>%
  summarise(count = n())

```

## Package: ggplot2

Ggplot2 has many functions for graphing different types graphs (e.g. histograms, box plots) with titles, labels, colors, and keys.

```{r ggplot}

ggplot(data=weather5, aes(x=how_check_weather, y=count)) +
  geom_bar(stat="identity", fill="darkblue", position = "dodge")+
  ggtitle("How Individuals Typically Check The Weather") +
   ylab("Frequency") + xlab("Method of checking weather") +
       theme(axis.text.x = element_text(angle=20, hjust=1))
  
```

## Conclusion

Tidyverse is a unique collection of packages useful for cleaning, tidying, manipulation, analysis, and visualization in a more efficient way - shown in my example above using a dataset about checking the weather.
