---
title: "Assignment Web APIs"
output: html_document
date: "2024-10-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

## Working with Purrr

In this sample vignette, I will walk through some of the functionalities of the the _purrr_ Tidyverse package using an [agriculture dataset](
https://www.kaggle.com/datasets/bhadramohit/agriculture-and-farming-dataset) from Kaggle.

According the [Tidyverse](https://purrr.tidyverse.org) site, purrr "enhances R’s functional programming (FP) toolkit by providing a complete and consistent set of tools for working with functions and vectors." It replaces a conventional "for loop" with a collection of functions centered around the _map_ function to introduce an easy-to-read, compact syntax, that is powerful and performant.

In this tutorial, I will go over the following functions
- map
- map_* family
- map2
- pmap

Before we get started, lets read in our csv file and prep the _farm_data_ dataframe.

```{r read-file}
#setwd("")
farm_data <- read.csv('https://raw.githubusercontent.com/mcastro64/d607-assignments/refs/heads/master/assignment-07/agriculture_dataset.csv') %>%
  rename_all(~str_to_lower(.)) %>%    # Convert all column names to lowercase
  rename_all(~str_replace_all(., "\\.", "_")) %>% # Replace spaces with underscore
  rename_all(~str_replace_all(., "_$", "")) # remove trailing _

glimpse(farm_data)
```

### Using _map_
The _map_ function makes a list by passing a list or an atomic verctor and a function as parameters in the format __map(.x, .f)__, where .x is our list/vector and .f is our function.

In example 1, we want to use Prrr to calculate the mean for the numerical columns of our sample dataframe. We first pass the dataframe to the __dplyr__ function _"select"_, then use _"where"_ to select the correct columns using _is.numeric_ to return only the columns with numeric values. We can then pass this list into the _map_ function using the .x parameter and pass function "mean" using the .f parameter. Example 2 shows an abbreviated version of the same calculation.

__Example 1__
```{r example-1}
ex1_list <- farm_data |>
  select(where(is.numeric)) 

map(.x = ex1_list, .f = mean)

```

__Example 2__
```{r example-2}
ex2_list <- farm_data |>
  select(where(is.numeric)) |>
  map(mean)

ex2_list
```
### Using other functions in the _map_*_ family

In addition to the _map_ function that returns a list, Purrr has dedicated functions that return specific vector types. Each functions named with the "map_" prefix followed by an abbreviation of the vector type, as seen in Example 3.

- map_lgl() returns a logical vector
- map_int() returns an integer vector
- map_dbl() returns a double vector
- map_chr() returns a character vector

__Example 3__
```{r map_dbl_example}

farm_area_miles <- map_dbl(farm_data$farm_area_acres, function(.x) {
  return(.x/640)
})

farm_area_miles
```

### Returning a Dataframe

Purr can also return a dataframe using the _map_df_ function. Example 4 modifies the _map_dbl_ example to return a dataframe of farm area in acres and miles. This example also uses an inline funciton notation whose variable __.x__ represents the rows of the data variable ("farm_data$farm_area_acres") that was passed into the _map_df_ functions. The function returns a data frame.

__Example 4__
```{r map_df_example}

map_area_df <- map_df(farm_data$farm_area_acres, function(.x) {
  return(data.frame(acres = .x, miles = .x/640))
})

map_area_df

```
### Handling multiple inputs

Purrr also has two functions for handling more than 1 list at a time:
- _map2_- applies a function to two elements
- _pmap_- applies a function to a list of elements

** Working with _map2_**

Example 4 shows how to work with _map2_. We pass two lists into variables .x and .y. These variables are then passed into the function when initializing it and manipulated within its body. The function returns a list.

__Example 4__
```{r map2-examples}

# map2 example
map2_ex <- map2(.x = farm_data$yield_tons, .y = farm_data$farm_area_acres, function(.x, .y) {
  return(.x /.y)
})

# print type and first 3 rows
class(map2_ex)
print(map2_ex[1:3])
```

**Working with _pmap_**

_pmap_ allows you to loop through more than two by passing in a list of those elements. Rather then use .x or .y, you need to specify the exact names of list elements in your function call. Example 5 generates randomized sample data with _pnorm_ using _pmap_. The first half of the code chunk is used to create a list for means (means), standard deviations (sds), and sample size (samplesize) from the farm_data dataframe. We then pass these variables into our function when calling _pmap_.

__Example 5__
```{r pmap-examples}

# construct list for this example
calcs_list <- farm_data |>
  group_by(crop_type) |>
  mutate(
    means = mean(yield_tons),
    sds = sd(yield_tons),
    samplesize = 50
  ) |>
  subset(select=c(means, sds, samplesize))

# pmap_example
set.seed(1234)
pmap_ex <- pmap(calcs_list, function(means, sds, samplesize) {
  data.frame(sample = rnorm(n=samplesize, mean=means, sd=sds))
})

ggplot(pmap_ex[[1]], aes(x=sample))+
  geom_histogram(bins = 50)

ggplot(pmap_ex[[2]], aes(x=sample))+
  geom_histogram(bins = 50)
```
### Nesting and combining map functions

The Purrr functions can also be used in combination with other Tidyverse packages and functions. Example 6 uses the _map_ function within _mutate_ to add a linear model using the _lm_ function to every "crop_type" in our farm data. Before we can pass the data to _map_, the example also uses _nest_ to convert the grouped dataframe into a list where the group is in the first column and the remaining columns from our original dataframe are "nested" into the second column of our list. The returned object adds the result of __lms__ as a third column, which can be further cleaned up and _unnest_'ed for other applications.

__Example 6__
```{r farm_data_lms}

farm_data_lms <- farm_data |> 
  group_by(crop_type) |> 
  nest() |>
  mutate(lms = map(data, ~lm(yield_tons ~ fertilizer_used_tons + pesticide_used_kg, data = .)))

farm_data_lms[[3]][1]

```

## Conclusion

The Purrr dataset is a great alternative to using _for_ loops. One advantage is more human-friendly syntax that compresses multiple lines of code into an easier to ready format. 

This tutorial only covers some of the most common functions of the _Purrr_ package, but the package includes many more promising tools to make your R programming more efficient, including _walk_, _modify_ and other _map_ variants like _imap_ and _map_if_. For more information, visit the [Purrr documentation](https://purrr.tidyverse.org/reference/map.html) or review the [Posit Purrr Cheatsheet](https://github.com/rstudio/cheatsheets/blob/main/purrr.pdf).
