---
title: "week 9 tidyver/github lab"
author: "Chi Hang(Philip) Cheung, updatd by Inna Yedzinovich and Yanyi Li"
date: "2024-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## R Markdown

Load data frame:
```{r cars}
url<- "https://raw.githubusercontent.com/stormwhale/data-mines/refs/heads/main/user_behavior_dataset.csv"
data<- read.csv(url)
```

String detect two key words: "pixel" and "galaxy".
data is piped into filter then combined with the function of str_detect to catch the desired character pattern.
```{r}

pi_gal<- data %>% 
  filter(str_detect(Device.Model,
                    regex("Pixel|Galaxy", ignore_case = TRUE)))
head(pi_gal)
```
To use str_extract to extract only numbers from a column mixed with characters:
"\\d+" is a regular expression for any digits
mutate is used to create a column to store the extracted numbers

```{r}
phone_gen<- data %>% 
  mutate(Phone_generation=str_extract(Device.Model, "\\d+")) %>% 
  select(c(Device.Model, Phone_generation))
head(phone_gen)
```
Now we extract phone generation and remove duplicates: 
```{r}
phone_gen <- phone_gen %>%
  mutate(Phone_generation = str_extract(Device.Model, "\\d+")) %>%
  group_by(Device.Model, Phone_generation) %>%
  summarise(Duplicate_Count = n(), .groups = 'drop')
head(phone_gen)

```

----------------------------------------------------------------------------------------------------------------------------
(Extended by Yanyi Li 11/26/2024)

We can use the ggplot2 package to create plots for the data frame above. One example is a bar graph for the duplicate counts.
```{r}
ggplot(data = phone_gen, aes(x = Device.Model, y = Duplicate_Count, fill = Device.Model)) +
  geom_bar(stat = "identity", width = 0.7) + 
  theme(axis.text.x = element_text(angle = 15))
```

The ggplot package is extremely useful for creating and customizing plots. In the graph above, I adjusted the bar width to 0.7 to create more space between the bars. Additionally, I assigned a different color to each bar to enhance the visual appeal. To improve the readability of long x-axis labels, I tilted them at a 15-degree angle.
