---
title: "Assignment_tidyverse_updated"
author: "Andreina A, extended by Marco Castro"
date: "2024-11-02"
output: html_document
---


```{r}
library(tidyverse)
```


**The data I used for this assignment was from fivethirtyeight and it's on the alcohol consumption from different countries in 2010. To load the data I used read.csv, since the file is a csv in github. Read.csv is from the readr library which is also in the tidyverse package. The glimpse function is from the `tibble` library from tidyverse, that is used to view the data set and get information on the amount of rows and columns. This data set has 193 rows and 5 columns.**


Article: https://fivethirtyeight.com/features/dear-mona-followup-where-do-people-drink-the-most-beer-wine-and-spirits/

Data: https://github.com/fivethirtyeight/data/tree/master/alcohol-consumption

```{r}
DF1 <- read.csv('https://raw.githubusercontent.com/fivethirtyeight/data/refs/heads/master/alcohol-consumption/drinks.csv')
glimpse(DF1)
```

From tidyverse's __tidyr__ library, the `pivot_longer` function can be used to reshape the data frame by collapsing four columns into two columns, resulting into a lengthen data frame.

```{r echo=FALSE}
DF1<-data.frame(DF1%>%
  pivot_longer(c(beer_servings, spirit_servings, wine_servings,total_litres_of_pure_alcohol), names_to = 'alcohol_type_of_serving', values_to = 'serving_amount'))
head(DF1)
```
**`drop_na` is a function from __tidyr__ that drops missing values and trues the values into floats.**

```{r echo=FALSE}
head(drop_na(DF1,serving_amount))
```


The `ggplot` library from tidyverse is used to plot data sets. We can use the `filter` function to only display data for serving amount that are greater than or equal to 347.

```{r ggplot-ex}
DF2 <- DF1 |>
  filter(serving_amount>=347)
ggplot(DF2, aes(x=country, y=serving_amount, fill=alcohol_type_of_serving)) +
  geom_col(position="dodge")
```

In addition to `filter` function, __dplyr__ contains many other helpful functions. In the example below, we used `group_by` to group our data by _alcohol_type_of_serving_, use `mutate`, another __dplyr__ function to add a column of the average serving amount per alcohol type, then `ungroup` to be able to use access the value in the new column from within our original rows. In the the example below, we use mutate again to add another column (_pct_above_avg_) for the number of times above the average serving amount for each country.

```{r group example}

DF3 <- DF1 |>
  filter(alcohol_type_of_serving != 'total_litres_of_pure_alcohol') |>
  group_by(alcohol_type_of_serving) |>
  mutate(
    avg_serving_amount = mean(serving_amount)
  ) |>
  ungroup() |>
  mutate(
    pct_above_avg = floor(serving_amount/avg_serving_amount),
  ) 
```

 We also used the __tidyr__ `separate_wider_delim` function to split the _alcohol_type_of_serving_ column into two columns, using "_" as a delimiter. We dropped the _right_ column created. Next, we used the __dplyr__ `select` function to select which columns we want to keep in our dataframe. In this example, we used the flag `-` to specify that we should drop the columns in the list. Finally, we filtered for only the countries that with serving amounts 4x's the overall average for each alcohol type then plotted using faceting by each alcohol type. 
 
```{r sep_wider_ex}
DF3 <- DF3 |>
  separate_wider_delim(alcohol_type_of_serving, delim = "_", names = c("alcohol_type", "right")) |>
  select(-c('right')) |>
  filter(pct_above_avg > 4) 
  
 
head(DF3)

ggplot(DF3, aes(y=country, x=serving_amount, fill=alcohol_type)) +
  geom_col(position="dodge") +
  facet_grid(vars(alcohol_type)) + 
  theme(legend.position = "none")
```

# Conclusion

**Tidyverse is a package that is handy for manipulating and transforming data. I was able to use filter to transform the data, glimpse to view the data, read.csv to read the data from a csv file into a data frame, pivot_longer to reshape the data, ggplot to show a visual of the data I was exploring.**