---
title: "DATA607_GITHUB CREATE"
author: "Ying Fang Lee"
date: "2024-12-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Introduction

In this assignment, you’ll practice collaborating around a code project with GitHub. You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

Using article found on Kaggle [“which movie should I watch today?”](https://www.kaggle.com/datasets/hassanelfattmi/which-movie-should-i-watch-today?resource=download) which included 4 csv files called movies, FilmDetails, MoreInfo, and PosterPath.csv. I will use the MoreInfo file to create a sample vignette that demonstrates how to use the Tidyverse packages.

#### Load data from Github

```{r}
# Load data from Kaggle as dataframe
url <- "https://raw.githubusercontent.com/amily52131/DATA607/refs/heads/main/Final_Project/Data/MoreInfo.csv"
MoreInfo <- read.csv(url)
```

#### From the tidyverse library we are using functions from the dplyr package

##### **mutate()**

The mutate function can be used two ways:

1.  Create a new column in data frame based on calculations from another column

2.  Update existing values in a column from data frame with logical operations

```{r}
# convert runtime to total_length by separating hours and minutes
MoreInfo <- MoreInfo %>% 
  mutate(hours = as.numeric(gsub("([0-9]+)h.*","\\1", runtime )),
         minutes = as.numeric(gsub(".* ([0-9]+) min.*","\\1", runtime)),
         total_length = hours * 60 + minutes
        )

# Update the budget and revenue column into integer 
MoreInfo <- MoreInfo %>% 
  mutate( budget = as.integer(gsub("[$,]","", budget)),
          revenue = as.integer(gsub("[$,]","", revenue))
  )
head(MoreInfo, 5)
```
##### **select()**

The select function lets us choose the columns based on condition or by column name

```{r}
# Only select the columns that are relevant
MoreInfo <- MoreInfo %>% 
  select(id, budget, revenue, total_length)
head(MoreInfo, 5)
```

#### **filter()**

The filter function lets us choose rows based on conditions

```{r}
# For this part we are removing observations where budget and revenue is NA
MoreInfo <- MoreInfo %>% 
  filter(!is.na(budget), !is.na(revenue))
head(MoreInfo, 5)
```

### From the tidyverse library we are using functions from the ggplot2 package

#### ggplot2 package

ggplot2 is used to plot graphs and it can be integrated with other packages in the tidyverse library

```{r}
# Plotting the budgets of movies
ggplot(data = MoreInfo, aes(x = budget)) +
  geom_histogram(bins = 30) + 
  scale_x_continuous(labels = function(x) x/1000000) + #scale by million
  labs(title = "Movie Budget", x = "Budget (million)", y = "Number of Movies") +
  theme_minimal()
# the most common budget for movies are all less than 100 million dollars 
```



















