---
title: "Assignment 9 - TidyVerse/GitHub"
author: "Zaneta Paulusova"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r}
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
```

### Instructions
In this assignment, you’ll practice collaborating around a code project with GitHub.  You could consider our collective work as building out a book of examples on how to use TidyVerse functions.

GitHub repository: https://github.com/pkowalchuk/FALL2024TIDYVERSE

Your task here is to Create an Example. Using one or more TidyVerse packages, and any dataset from fivethirtyeight.com or Kaggle, create a programming sample “vignette” that demonstrates how to use one or more of the capabilities of the selected TidyVerse package with your selected dataset. (25 points)


The dataset that I used was from Kaggle on Marvel Movies Ratings from multipe diffrent sites and scoring ranges. I will be only looing on the one that the score is out of 100. (https://www.kaggle.com/datasets/sarthakbharad/marvel-movies-dataset?select=Marvel_Movies_Dataset.csv).

### Load the dataset from GitHub
```{r}
marvel_movies_data <- read.csv("https://raw.githubusercontent.com/ZanetaP02/DATA-607/refs/heads/main/Marvel_Movies_Dataset.csv")
colnames(marvel_movies_data)
```
### Rename columns
```{r}
mm_data<- marvel_movies_data

names(marvel_movies_data)[3] <-"Director1"
names(marvel_movies_data)[4] <-"Director2"
names(marvel_movies_data)[5] <-"Release_Date"
names(marvel_movies_data)[6] <-"IMDB_out_of_10"
names(marvel_movies_data)[7] <-"IMDB_Merascore"
names(marvel_movies_data)[8] <-"Rotten_Tomatoes_Critics"
names(marvel_movies_data)[9] <-"Rotten_Tomatoes_Audience"
names(marvel_movies_data)[10] <-"Letterboxd_out_of_5"
names(marvel_movies_data)[11] <-"Cinema_Grade"
names(marvel_movies_data)[12] <-"Budget_in_Million"
names(marvel_movies_data)[13] <-"Domestic_Gross_in_Million"
names(marvel_movies_data)[14] <-"Worldwide_Gross_in_Million"
```

### Print column names to verify
```{r}
print(colnames(mm_data))
```
### Summary of the data
```{r}
summary(mm_data)
```

### Moving the sourse that are out of 100 to one column
```{r}
m_data <- mm_data %>%
  pivot_longer(c('IMDB_Merascore', 'Rotten_Tomatoes_Critics', 'Rotten_Tomatoes_Audience'), names_to = "Rated_By", values_to = "Scores")
m_data
```
### Graph of the three different out of 100 Scorers
```{r}
ggplot((m_data), aes(x = Title, y = Scores, fill = Rated_By)) + geom_col(position = position_dodge())+
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 70, hjust = 1))
```

### Conclution
From the graph it it show that under Rotten Tomatoes score from Critics or Audience it looks like the rating are higher then underIMDB Merascore. there are other scores, but they would need to be converted to be out of 100 to see more of a range of scores through different Rating.